### 2013-06-29 rebuttal.txt
Updated the rebuttal and fixed typo in the article.

We would like to thank the reviewers for the valuable feedback.

Generality:

Barth et al. show how a carefully crafted file of a benign format can manipulate the
browser's content-sniffing algorithm into rendering the file as malicious HTML. With this
and other similar attacks as a starting point, we show that the issue extends beyond
content sniffing and demonstrate how other formats can be abused in a web context, thereby
generalizing polyglot-based attacks.

Following the suggestion, we plan to make the claims of generality more precise, with our
high-level presentation of polyglots and their abuse possibilities as general, and the
focus on new PDF-based attacks as particular. Similarly, we agree it is a good idea to
clarify the general vs. particular aspects of mitigation.

Novelty:

The novelty of this paper is twofold: generalizing polyglot attacks and showing a practical 
example of a vulnerable format. As mentioned, malicious PDFs or Flash is no surprise. However, we 
show how the flexibility of the PDF format accommodates hiding such	a malicious PDF document in a 
benign format. We also discuss PDF as a delivery mechanism for an embedded malicious Flash. The 
result has the flexibility of PDF to be crossbred with other formats, and the capabilities of PDF 
and Flash combined.

As for other vulnerable formats, future work mentions scriptable media formats, e.g., Advanced 
Streaming Format (ASF), as another problem area. However, space constraints prohibits us from 
further exploring these formats within the scope of this paper.

Advanced mitigation:

Regarding more advanced defenses, we will be happy to include in the paper a discussion of
most promising directions in this space. For example, note that when content is loaded in
a specific context, as it is in the case of the paper, the browser already has exact
knowledge as to what type it will interpret the requested content as. A natural defense
technique can be to send the expected type along with the request. The web server then
can compare the types and react accordingly. Further, it can be implemented in the web
server itself (as opposed to the web application), since the web server makes the final
decision on the supplied content-type. This co-operation between client and server can help 
mitigate both syntax injection and content smuggling. A proof-of-concept implementation would 
require minimal adjustments on both client- and server-side and an evaluation could be done easily.

The evaluation may well be too focused on PDF-based polyglot attacks, and would benefit from more 
general tests. Evaluation of mobile browsers would indeed make an interesting addition, and 
certainly one we would like to delve into further.

Concerning XML External Entities (XXE) in Adobe Reader, the problem of XXE was previously known 
(since ca. 2005), but considered to be of low priority. Our reports may have influenced Adobe's 
decision to finally remove them, but it is not known to us to what extent.

As suggested, we plan to make the treatment of the related work more coherent. Also, we
are happy to follow the suggestion to update the visualization with a more professional
appearance and to move the listings to their appropriate positions.

----------------------- REVIEW 1 ---------------------
PAPER: 496
TITLE: Polyglots: Crossing Origins by Crossing Formats
AUTHORS: (anonymous)

OVERALL EVALUATION: 2 (accept)

----------- REVIEW -----------
This paper examines an attack based on interpretation of scripts. The idea is that a valid script in language A is also valid in language B but does something very different. So, a script that will pass checks for maliciousness in language A is uploaded in such a way that, when a client downloads it, the script is interpreted as language B -- and that produces a malicious action.

This is not a new idea. The authors point out several examples, including GIFAR and a cross-origin CSS attack. Another good example is a type flaw in cryptographic protocols (see [1]). This paper analyzes the format of PDF objects in considerable detail, and demonstrates how to use that format to launch this type of attack.

[1] C. Boyd, "Hidden assumptions in cryptographic protocols'", IEE Proceedings E: Computers and Digital Techniques 137(6) pp. 433-436 (Nov. 1990);

The paper is an excellent, in depth study of PDFs, and how to launch this type of attack on the web. One improvement that the authors should make is to tone down their initial claim that it is a "generalization and systematic study of polyglot attacks". While it does discuss elements of such attacks in general, the discussion is not at the depth this claim suggests. The discussion of how to do this type of attack in the context of the web is general, but that does not make the paper a general and systematic study of this type of attack.

### Jonas ###
# To Andrei: Say something clever about being systematic, or take it back?
#############

Some minor comments.

Section 4 begins with a sentence that says "investigate the persistence" -- you mean "investigate the prevalence", because persistence means the vulnerabilities "go on resolutely or stubbornly in spite of opposition, importunity, or warning".

### Jonas ###
# To Andrei: Fixed.
#############

Some typographical errors:

Section 3.1.2, line 2: "indirect objects" should be "an indirect object"
Section 3.1.2, subsection "Stream objects", line 2: "shall always be" should be "is", to preserve the parallelism of the sentence Section 5.1.2, paragraph 2, line 4: "that is" should be "that are"

### Jonas ###
# To Andrei: All fixed.
#############

----------------------- REVIEW 2 ---------------------
PAPER: 496
TITLE: Polyglots: Crossing Origins by Crossing Formats
AUTHORS: (anonymous)

OVERALL EVALUATION: 1 (weak accept)

----------- REVIEW -----------
This paper proposes several new web attacks based on the idea that a file format may appear different to the webserver than the end user.
Such programs can be abused to create insecure communications across Internet origins.  The authors notice that their attacks are additional instances of the same general problem, with one notable example being the GIFAR GIF+Java Archive attack.  The authors propose that their attacks and others be considered a general category, called polyglots. Their evaluation claims 5 of the top 100 web sites are vulnerable.  The authors analyze their new polyglots, and propose mitigations.

The overall attacks are interesting and fun to read. I like the idea of generalizing the notion of specific previous instances. I think this paper will result research in more generalized defenses.

The mitigation measures aren't that exciting. For example, the authors say in 5.1.1 that there isn't a general purpose answer for syntax injection attacks. It would be nice to have thought more deeply about more general mitigation measures. I agree that there will not be a complete defense, but the authors stop at the obvious.  There is a set of security people who believe attack papers without proposed defenses aren't of much value. I don't side with them, but certainly understand that position. Having more in-depth defenses would strengthen the paper against such attacks.

### Jonas ###
# To Andrei: This is where I would have liked to add the new mitigation idea.
#
# 
#############

The basic mechanics of grammar and spelling are fine. (small exception I just noticed: "While there are exceptions to this policy, e.g.". You know the problem....there is a missing comma.)

### Jonas ###
# To Andrei: Fixed.
#############

However, I found several sections unclear and hard to read.  For example, "According to the SOP, content loaded in a plug-in is considered to belong to the origin from which it was loaded, thus allowing the polyglot to communicate with the vulnerable origin."  The word "it" could refer to multiple nouns.

### Jonas ###
# To Andrei: Corrected.
#############

"will activate the malicious polyglot".  This isn't clear; what does it mean to activate a polyglot?  Given specific names might be one way to solve the problem. In a sense, you have program M that is malicious wrapped in program N that is not.  It seems like the vulnerable website seems N||M, and interprets this as type N, while the victim sees M, the malware.

### Jonas ###
# To Andrei: Not corrected yet, thinking about it.
#############

I always find that cartoon diagrams, such as in Figure 1, look unprofessional.  This isn't a dig saying the paper isn't professional...it is. Just that the visual aids look hokey and unappealing.  A simple fix is to remove the cartoon and just use block diagrams.  Also, pay attention to scaling; when printed out the diagram is currently blurry.

### Jonas ###
# As is suggested, the visualization will be updated for a more professional appearance, 
#############

Move the listings that appear at the end of the paper in place. Not every listing has to be a float; you can just put it right underneath where you want to talk about it.

### Jonas ###
# and the listings moved to their appropriate positions.
#############

----------------------- REVIEW 3 ---------------------
PAPER: 496
TITLE: Polyglots: Crossing Origins by Crossing Formats
AUTHORS: (anonymous)

OVERALL EVALUATION: 0 (borderline paper)

----------- REVIEW -----------
This paper describes the concept of polyglots, programs that are syntactically valid in multiple languages. The focus of this work is on malicious polyglots, those programs that act benignly when loaded as one language in a particular context, but can be interpreted in a malicious manner by another language or browser plug-in. Particularly with plug-ins, it is possible to circumvent protections such as the same-origin policy, and these format-crossing polyglots can potentially inject syntax or smuggle undesired content in with them, which can then be interpreted and executed.
An example content smuggling scenario is described, as are the consequences (e.g., CSRFs and cross-origin leakage). The authors find that PDFs are particularly well suited to be polyglots, as a Flash to a slightly lesser extent. An evaluation shows that every major browser has some susceptibility to these attacks, with Chrome slightly better when it come to defending against malicious PDFs. Mitigation strategies at the server and client sides are considered.

This paper brings up an interesting area of focus, and explains it well. The paper was enjoyable to read and for someone who has not read considerably into the recent web literature, this was a very good overview of attacks that these polyglots can cause.

However, the novelty of these attacks is somewhat limited. It is hardly surprising that PDFs 
can be designed in a malicious fashion - Acrobat is well known for vulnerability to not only 
being taken over through buffer overflows, but for executing malicious PDFs as well. This was 
discussed in some length at BlackHat last year, where an example of a PDF hiding malicious 
JavaScript was shown by Jason Jones ("State of Web Exploit Kits").

### Jonas ###
# The novelty of this paper, in relation to PDFs, lies not in showing how a PDF can be malicious, 
# but rather to show how the flexibility of the format makes it ideal for hiding a malicious PDF 
# in another format.
#############

Similarly, it should come as a surprise to nobody to read that Flash is also a vulnerable format.

### Jonas ###
# In relation to Flash, it is not attempting to show that Flash is vulnerable, but rather to show
# how PDF can serve as a delivery mechanism for malicious Flash.
#############

It would have been interesting to delve into more surprising formats. The paper from Barth et al. 
looks at malicious PostScript, and PDF, while not being quite as bad, has certainly suffered from 
its share of problems.

### Jonas ###
# Future work mentions scriptable media formats, such as Windows Media Format, as another problem area.
# However, space constraints prohibited us from further exploring these formats within the scope of this paper.
#############

The evaluation for this paper felt somewhat pro forma. I would have rather seen a more extensive 
set of tests against the individual browsers,

### Jonas ###
# THe evaluation is indeed focused on PDF-based polyglots, thereby not making it general 
#############

and perhaps to have considered mobile web browsers 
as well.

### Jonas ###
# Mobile browsers would indeed make an interesting addition.
#############

Where I think the paper was particularly lacking was on the mitigation side. For a venue such as 
this, I would have liked to have seen an implementation of one of the mitigation solutions and a 
full evaluation of security and performance. This could be done either on the browser or server 
side. Without that, this paper has more of an informational feel to it.

### Jonas ###
# This mitigation technique 
# Implementing this mitigation technique would require adjustments on both
#
#############

Smaller issues:

- The related work reads like a laundry list of papers, and could stand to be more coherent in its treatment 
of other work, thus better positioning what the novel contribution of this paper is. The authors claim a 
generalization of the polyglot idea, but there were not enough cases to show a large generalization beyond 
what Barth et al. had already shown.

### Jonas ###
# To Andrei: Don't know how to tackle the "laundry list" part. Remove some related work?
#
# Barth et al. only consider formats that are combined with malicious HTML. We show that the issue 
# extends far beyond HTML, and that other formats can be abused in a web context, thereby making the 
# results more general.
#############

- Did the authors influence the design of XML External Entities as appears
  to be suggested in Section 3.4? If so, that is certainly worth playing up
  as a contribution.

### Jonas ###
# The problem of XML External Entities was previously known (since 2008), but considered to 
# be of little importance. Our reports may have influenced Adobe's decision to finally remove them,
# but it is not known to us to which extent.
#############